Splunk Tutorial
Next »
Venerated for real-time Operational Intelligence and secure data processing, Splunk offers the best platform for log analytics. The company has achieved the top position for developing a phenomenal open-source log analysis software until now. Exceptionally strong in dealing with today’s large volumes of data , Splunk provides acute efficiency to search, analyze, store and process data.

Some of the special features of Splunk are:

Stores data in its index and therefore separate database required
Discovers useful information automatically without searching manually
Converts log data into Visual graphs and reports to simplify analysis, reporting and troubleshooting
Get Splunk Certification in just 12 Hours
GET CERTIFIED
This Splunk free tutorial is an introduction to the powerful tool to work with machine generated data. This tutorial for beginners includes architecture, configuration, knowledge objects, Splunk files, indexer,forwarder and more. Enterprises find it a powerful and influential market leader offering innovative solutions to access machine data. Thousands of customers around the globe are implementing Splunk modules to make their businesses more secure and profitable.

Splunk versus ELK

Though ELK (ElasticSearch, Logstash and Kibana) is a strong competitor of Splunk, but there are certain situations when Splunk outshines ELK. Read below to find out:

Criteria	Splunk	ELK (ElasticSearch, Logstash and Kibana)
Feature	Log management and analysis platform	Consolidated data analytics platform
Parsing	Parses when searches are executed	Parses when data is ingested
Ease of Use	More features	Lesser features
Search capabilities	Wide array of options	Limited search capability
Enterprises find it a powerful and influential market leader offering innovative solutions to access machine data. Thousands of customers around the globe are implementing Splunk modules to make their businesses more secure and profitable.

You can be a valuable part of those hundreds of passionate users applying Splunk solutions on their data by acquiring conceptual as well as advanced-module knowledge of Splunk components via this tutorial. Having a forte in training individuals and professional worldwide, we believe in enlightening you with major topics on the technology you wish to adopt.

Here is a Splunk tutorial to gain a competitive edge over peers and co-employees as you can promptly answer the related questions, provide spontaneous solutions to managing big streams of data and monitor and secure data regardless of their users and sources.

Free Ebook: Step by Step Guide to Master Splunk
GET EBOOK
Since security of data is a critical aspect of every organization, what can be better than mastering data security using a modern big data system? This tutorial is a complete learning kit for beginners as well as advanced level professionals, which can enable you to take wise business decisions related to your data. From Splunk Installation to searching to grouping events, we train you on core technical concepts of this big data platform.

The biggest challenge with big data is effective searching and here is the quick guide navigating you through proper searching channels such that you master the text-in-search techniques using Splunk.
A detailed operational understanding on Search Processing Language (SPL) with examples is a great learning benefit you can achieve.
The tutorial covers quick and easy-to-understand tips and tricks to enrich your data for processing purposes.
Monitoring, Alerting, Grouping events and gaining expert-level understanding on Lookup tables are some of the excellent coaching topics of this tutorial.
The refinement and variation in which these topics are explained are flawless. You won’t require to reread or go over it multiple times – it is easy-to-comprehend and implement.

Learn more about Splunk in this insightful blog now!
